On verifying AhirV2 generated VHDL using software testbenches

### 2014-09-02 Documents/AhirV2/AhirV2.tex
Get the documentation into better shape.

AhirV2 is a set of tools 
\begin{itemize}
\item which can convert a C/C++ progam to  an equivalent hardware
implementation (described in VHDL).
\item which provide an algorithmic approach to the design of 
hardware.
\end{itemize}

@@

\begin{itemize}
\item Given a high-level C/C++ program, we rely on
an LLVM (www.llvm.org) compatible compiler such as clang (www.clang.org)
or dragon-egg (dragonegg.llvm.org) to produce LLVM byte code.
Currently, the AhirV2 flow uses LLVM byte code as a starting point.
\item The LLVM byte-code program is compiled
to an intermediate assembly form.  AhirV2
introduces an intermediate assembly language
\Aa  which serves as a target for sequential
programming languages (such as C/C++) as
well as for parallel programming languages.
An \Aa program consists of modules (analogous
to sub-programs in C/C++) which can call each 
other, and can communicate through storage objects
as well as through pipes (first-in-first-out buffers).
\item From the \Aa description, a virtual
circuit (described in a virtual circuit
description language \vC) is generated.
The chief optimization carried out at this
step is the decomposition of the system memory
into disjoint spaces based on usage
analysis (this considerably improves the
available memory bandwidth and reduces
system cost).  A \vC description also
consists of modules: however, the modules
are presented in a factored form (control X data X storage).
\item From the \vC description, a
VHDL description of the system is generated.
The system consists of modules, memory spaces
and FIFO buffers.  The modules are further
broken down into a control-path (a live and safe
Petri net), and a data-path (a graph of operators
and wires).
The chief optimization carried out at this
stage is resource sharing.  The \vC description
is analyzed to identify operations which
cannot be concurrently active and this information
is used to reduce the hardware required.
\item 
The VHDL description produced from 
\vC is in terms of a library
of VHDL design units which has been developed
as part of the AhirV2  effort.  This library
consists of control-flow elements, data-path
elements and memory elements.
\end{itemize}

Thus, to generate hardware using the AhirV2 flow,
it is possible to start at C/C++, at the \Aa level
at the \vC level or at the VHDL level (or a combination
of all these levels).  Starting at a higher level is
easier for the programmer, but using lower level representations
will probably lead to more efficient hardware.
Typically, if we start from C/C++, circuits
produced by the AhirV2 flow are upto two orders of
magnitude more energy-efficient than
a processor \cite{ref:dsd2010}.

Currently, there are only two restrictions in mapping
a C/C++ program to VHDL using the AhirV2 flow:
\begin{itemize}
\item No recursion, no cycles in the call-graph of the original
program.
\item No function pointers.
\end{itemize}

@@

void bar()
{
   while(1)
   {
     uint32_t data = read_uint32("midpipe");
     write_uint32("outpipe", ~data);
   }
}
\end{verbatim}
In this example, the read/write functions are assumed
to use named pipes for inter-process communication.

@@

This tool takes LLVM byte code and converts it into an
\Aa file.
\begin{verbatim}
llvm2aa [-modules=<listfile>] [-storageinit] bytecode.o > bytecode.aa
\end{verbatim}
The generated \Aa code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success, the tool returns 0.

The options:
\begin{itemize}
\item {\bf -modules=listfile} : Specify the list of functions in the bytecode
which should be converted to \Aa.   The names of these functions should be
listed in the text-file listfile. If absent, all functions
are converted.
\item {\bf -storageinit} :  Storage objects in the llvm bytecode
are explicitly initialized in the generated \Aa code.   An initializer
routine {\bf global\_storage\_initializer\_} is instantiated in
the \Aa code for this purpose.
\end{itemize}

@@

This linker tool takes a list of \Aa files, elaborates the program,
creates a global storage initializer, and
does memory space decomposition.  The externally visible memory space is
linked in one of two ways: either it is assumed to be external
and all accesses to it are routed out of the \Aa program,
or it is assumed to be internal and assumed to consist of
a memory object (an array of bytes).  External pointer dereferences
are handled as if they are directed at this memory object.
\begin{verbatim}
AaLinkExtMem [-I n] [-E obj-name] bytecode.o > bytecode.aa
\end{verbatim}
The generated \Aa code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success, the tool returns 0.

@@

@@

This tool takes a list of \Aa programs and converts them
to a \vC description. 
\begin{verbatim}
Aa2VC [-O] [-C] [-I obj-name] file1.aa file2.aa ...  > result.vc
\end{verbatim}
The generated \vC code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success the tool returns 0.

The options:
\begin{itemize}
\item {\bf -O} : if used, sequential statement blocks are parallelized
by doing dependency analysis.
\item {\bf -C} : if used, a C stub is created for every module that
is not called from within the system.  These stubs can be used to
interface to a VHDL simulator (or even drive hardware) to verify
the VHDL code generated by downstream tools.
\item {\bf -I obj-name} : if specified, all external memory references
are considered as being directed at the storage object named obj-name.
If not specified, then the tool will throw an error if it finds
a pointer dereference that cannot be resolved as pointing only to
storage objects declared inside the \Aa program.
\end{itemize}

@@

Takes a collection of \vC descriptions and converts them to
an AHIR system described in VHDL.
\begin{verbatim}
vc2vhdl [-O] [-C] [-a] [-e <entity-name] [-w]\
         -t/-T foo [-t/-T bar -t/-T bar2 ...]\
         -f file1.vc -f file2.vc ...  > system.vhdl
\end{verbatim}

The options:
\begin{itemize}
\item {\bf -t} : to specify the modules which are to be 
accessible from the ports of the generated VHDL system.
Such modules have to be top-level (that is, they cannot
be called from within the program).
Multiple top-level modules can be specified in this way.
The control and argument ports for these modules are
visible at the interface of the generated AHIR system.
\item {\bf -T} : to specify top-level modules which are to be 
free-running inside the AHIR system.
Multiple top-level modules can be specified in this way.
Such modules do not have any arguments and do not return
any values.  Their only mechanism of communication with
the world outside the AHIR system is through pipes.
The control ports for these modules are
{\bf not visible} at the interface of the generated AHIR system.
In the AHIR system, these modules are started on reset
and are run forever (restarted after they finish, forever).
\item {\bf -f} : specifies the \vC files to be analyzed. 
Multiple \vC files may be specified.  An object must be
defined before it is used, so the \vC files must be 
specified in the correct order.
\item {\bf -O} : optimize the generated VHDL by compacting
the control-path.  This does not change the resulting
hardware, but makes the generated VHDL file smaller.
\item {\bf -C} : the VHDL code has a system test bench which
interfaces to foreign code using a VHPI/Modelsim-FLI interface.
If this is not specified, the  generated test bench simply
instantiates the system and starts all top-level modules
off (you will need to fill in your own test bench here).
The C testbench is usually easier to write (it probably
already exists in the form of the original program).
\item {\bf -a} : try to minimize the area of the resulting
VHDL by sharing operators to the maximum extent possible
(allowing potential contention for resources).  This will
result in a slower (usually by 2X) system, but will
also reduce the area (usually by 0.5X).  If not specified,
two operations will be mapped to the same
operator  only if it can be proved that they cannot be active simultaneously.
\item {\bf -e top-entity-name} : The generated top-level VHDL entity
corresponding to the AHIR system is named top-entity-name.  The default
is test\_system.
\item {\bf -w} :  If specified, the VHDL system and test-bench are
generated as separate unformatted VHDL files.  You will need to
format these using the vhdlFormat command.
\item {\bf -s ghdl/modelsim} :  If {\bf ghdl} is specified
with the -s option, then the generated testbench (if -C is specified)
uses the VHPI interface to link with foreign code.  Otherwise,
the generated testbench (if -C is specified) uses the Modelsim FLI
interface to link with foreign code.
\end{itemize}

>>>>

AhirV2 is a set of tools which can convert a C 
description of a system to  an equivalent hardware
implementation (described in VHDL).  Using these tools,
it is possible to take an algorithmic approach to 
the design of  hardware.

@@

\begin{itemize}
\item Given a high-level C, we rely on
an LLVM (www.llvm.org) compatible compiler such as clang (www.clang.org)
to produce LLVM byte code.
Currently, the AhirV2 flow uses LLVM byte code as a starting point.
\item The LLVM byte-code program is compiled
to an intermediate assembly form.  AhirV2
introduces an intermediate assembly language
\Aa  which serves as a target for sequential
programming languages (such as C) as
well as for parallel programming languages.
An \Aa program consists of modules (analogous
to sub-programs in C) which can call each 
other, and can communicate through storage objects
as well as through pipes (first-in-first-out buffers).
\item From the \Aa description, a virtual
circuit (described in a virtual circuit
description language \vC) is generated.
The chief optimizations carried out at this
step are dependency based operation
ordering, dynamic loop-pipelining and decomposition of the system memory
into disjoint spaces based on static pointer
analysis (this considerably improves the
available memory bandwidth and reduces
system cost).  A \vC description also
consists of modules: however, the modules
are presented in a factored form (control X data X storage).
\item From the \vC description, a
VHDL description of the system is generated.
The system consists of modules, memory spaces
and FIFO buffers.  The modules are further
broken down into a control-path (a live and safe
Petri net), and a data-path (a graph of operators
and wires).
The chief optimization carried out at this
stage is resource sharing.  The \vC description
is analyzed to identify operations which
cannot be concurrently active and this information
is used to reduce the hardware required.
\item 
The VHDL description produced from 
\vC is in terms of a library
of VHDL design units which has been developed
as part of the AhirV2  effort.  This library
consists of control-flow elements, data-path
elements and memory elements.
\end{itemize}

Thus, to generate hardware using the AhirV2 flow,
it is possible to start at the C-level, at the \Aa level
at the \vC level or at the VHDL level (or a combination
of all these levels).  Starting at a higher level is
easier for the programmer, but using lower level representations
will usually lead to more efficient hardware.
Typically, if we start from C, circuits
produced by the AhirV2 flow are upto two orders of
magnitude more energy-efficient than
a processor \cite{ref:dsd2010}.

Currently, there are only two restrictions in mapping
a C program to VHDL using the AhirV2 flow:
\begin{itemize}
\item No recursion, no cycles in the call-graph of the original
program.
\item No function pointers.
\end{itemize}

@@

void bar()
{
   while(1)
   {
     uint32_t data = read_uint32("midpipe");
     write_uint32("outpipe", ~data);
   }
}
\end{verbatim}
In this example, the read/write functions are provided
as part of a {\em pipeHandler} library which is bundled with
the AhirV2 distribution.

@@

This tool takes LLVM byte code and converts it into an
\Aa file.
\begin{verbatim}
llvm2aa options bytecode.o > bytecode.aa
\end{verbatim}
The generated \Aa code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success, the tool returns 0.

The options:
\begin{itemize}
\item {\bf -modules=listfile} : Specify the list of functions in the bytecode
which should be converted to \Aa.   The names of these functions should be
listed in the text-file listfile. If absent, all functions
are converted.
\item {\bf -storageinit} :  Storage objects in the llvm bytecode
are explicitly initialized in the generated \Aa code.   An initializer
routine named
\begin{verbatim}
    global_storage_initializer
\end{verbatim}
is instantiated in
the \Aa code for this purpose.
\item {\bf -pipedepths=filename} : Specifies a file which contains
the depths of pipes which are part of the generated \Aa code.
\item {\bf -extract\_do\_while} : Innermost loops which are marked
using a call to the special function 
\begin{verbatim}
    _loop_pipelining_on_
\end{verbatim}
are extracted as pipelined do-while loops.  This is necessary
for automatic parallelization of inner loops in the 
generated hardware (substantial performance benefits can be
realized).
\end{itemize}

@@

This linker tool takes a list of \Aa files, elaborates the program,
creates a global storage initializer, and
does memory space decomposition.  The externally visible memory space is
linked in one of two ways: either it is assumed to be external
and all accesses to it are routed out of the \Aa program,
or it is assumed to be internal and assumed to consist of
a memory object (an array of bytes).  External pointer dereferences
are handled as if they are directed at this memory object.
\begin{verbatim}
AaLinkExtMem options file1.aa file2.aa ...  > linked.aa
\end{verbatim}
The generated \Aa code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success, the tool returns 0.

@@

\subsection{{\bf AaOpt}}

The optimization utility {\bf AaOpt} takes an \Aa program (list of
\Aa files) and produces an optimized version of the source program.
\begin{verbatim}
AaOpt options file1.aa file2.aa ... > optimized.aa
\end{verbatim}
The optimized \Aa code is printed to {\bf stdout}.  On success, 
the tool returns a 0 (else a non-zero).  Macro and inlined
function calls in the source code are substituted in place
in the optimized code.

The options:
\begin{itemize}
\item {\bf -r module-name} (optional) : specifies a root module in the
system.  Multiple root modules can be specified.  All dead code (which
is not reachable from a root module) is eliminated.
\item {\bf -I extmem-object} (optional) : similar to AaLinkExtMem,
this option specifies the name of the extmem-object in the
source \Aa files.
\item {-B} (optional) : if specified, add buffering to balance pipelined
loops so that loop performance is not bottlenecked by inadequate
buffering.
\end{itemize}

@@

This tool takes a list of \Aa programs and converts them
to a \vC description. 
\begin{verbatim}
Aa2VC options file1.aa file2.aa ...  > result.vc
\end{verbatim}
The generated \vC code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success the tool returns 0.

The options:
\begin{itemize}
\item {\bf -O} : if used, sequential statement blocks are parallelized
by doing dependency analysis.
\item {\bf -C} : if used, a C stub is created for every module that
is not called from within the system.  These stubs can be used to
interface to a VHDL simulator (or even drive hardware) to verify
the VHDL code generated by downstream tools.
\item {\bf -U} : memory subsystems will be unordered (that is,
will not guarantee in-order completion of accesses).  This
leads to a simpler memory subsystem, but more conservative
control flow.  The default is that all memory subsystems
are ordered (will complete read/write requests in the order that
they are accepted).
\item {\bf -r root-module} (optional): specifies a root module.
Code which is not accessible from a root-module is considered
as dead code and is ignored.
\item {\bf -I obj-name} : if specified, all external memory references
are considered as being directed at the storage object named obj-name.
If not specified, then the tool will throw an error if it finds
a pointer dereference that cannot be resolved as pointing only to
storage objects declared inside the \Aa program.
\end{itemize}

@@

Takes a collection of \vC descriptions and converts them to
an AHIR system described in VHDL.
\begin{verbatim}
vc2vhdl [-O] [-C] [-q] [-a] [-e <entity-name] [-w]\
         -t/-T foo [-t/-T bar -t/-T bar2 ...]\
         -f file1.vc -f file2.vc ...  > system.vhdl
\end{verbatim}

The options:
\begin{itemize}
\item {\bf -t} : to specify the modules which are to be 
accessible from the ports of the generated VHDL system.
Such modules have to be top-level (that is, they cannot
be called from within the program).
Multiple top-level modules can be specified in this way.
The control and argument ports for these modules are
visible at the interface of the generated AHIR system.
\item {\bf -T} : to specify top-level modules which are to be 
free-running inside the AHIR system.
Multiple top-level modules can be specified in this way.
Such modules do not have any arguments and do not return
any values.  Their only mechanism of communication with
the world outside the AHIR system is through pipes.
The control ports for these modules are
{\bf not visible} at the interface of the generated AHIR system.
In the AHIR system, these modules are started on reset
and are run forever (restarted after they finish, forever).
\item {\bf -f file-name} : specifies the \vC files to be analyzed. 
Multiple \vC files may be specified.  An object must be
defined before it is used, so the \vC files must be 
specified in the correct order.
\item {\bf -O} : optimize the generated VHDL by compacting
the control-path.  This does not change the resulting
hardware, but makes the generated VHDL file smaller.
\item {\bf -C} : the VHDL code has a system test bench which
interfaces to foreign code using a VHPI/Modelsim-FLI interface.
If this is not specified, the  generated test bench simply
instantiates the system and starts all top-level modules
off (you will need to fill in your own test bench here).
The C testbench is usually easier to write (it probably
already exists in the form of the original program).
\item {\bf -a} : try to minimize the area of the resulting
VHDL by sharing operators to the maximum extent possible
(allowing potential contention for resources).  This will
result in a slower (usually by 2X) system, but will
also reduce the area (usually by 0.5X).  If not specified,
two operations will be mapped to the same
operator  only if it can be proved that they cannot be active simultaneously.
\item {\bf -q } : if specified, do aggressive register insertion
to minimize the clock period.
\item {\bf -S bypass-stride } : by specifying the bypass stride
(an integer $\geq 1$),
the user can trade-off clock cycles versus clock period.  
The lowest clock period will be obtained for -S 1.   
\item {\bf -e top-entity-name} : The generated top-level VHDL entity
corresponding to the AHIR system is named top-entity-name.  The default
is test\_system.
\item {-L function-library} : AhirV2 provides some built in operator
functions which can be called from your code.  These are organized
as function libraries and this option specifies a function library
to look into when generating VHDL.  For example {\em -L fpu} gives
access to the floating point library which provides some useful
built in functions (e.g. fpalu32, fpalu64 etc.).
\item {\bf -w} :  If specified, the VHDL system and test-bench are
generated as separate unformatted VHDL files.  You will need to
format these using the vhdlFormat command.
\item {\bf -s ghdl/modelsim} :  If {\bf ghdl} is specified
with the -s option, then the generated testbench (if -C is specified)
uses the VHPI interface to link with foreign code.  Otherwise,
the generated testbench (if -C is specified) uses the Modelsim FLI
interface to link with foreign code.
\end{itemize}

### 2014-09-02 v2/CtestBench/docs/CtestBench.tex
Get the documentation into better shape.

Consider the following program (lets say it is in file ``prog.c''):
\begin{verbatim}
#include <stdlib.h>
#include <stdint.h>
#include <stdio.h>
#include <iolib.h>

uint32_t sum;

void set_sum(uint32_t x)
{
    sum = x;
}

uint32_t get_sum()
{
    return(sum);
}

void accumulate()
{
    while(1)
    {
        int nxt = read_uint32("in_data");
#ifdef SW
        printf("read %u\n", nxt);
#endif
        sum = (sum + nxt);
        write_uint32("out_data",sum);
#ifdef SW
        printf("wrote %u\n", sum);
#endif
    }
}

@@

Now to test this program, we can write a test-bench
such as this one (lets call this file ``testbench.c'').
\begin{verbatim}
#include <pthread.h>
#include <signal.h>
#include <stdio.h>
#include <stdlib.h>

#ifdef SW 
// for the read_*/write_* methods
#include <iolib.h>
#include "prog.h"
#else
// includes for the read_*/write_* methods
// as well as stubs for functions moved
// to hardware.
#include "vhdlCStubs.h"
#endif

@@

void *accumulate_(void* fargs)
{
   accumulate();
}

void *write_pipe_(void* a)
{
   write_uint32_n("in_data",(uint32_t*)a, 10);
}

void *read_pipe_(void* a)
{
   read_uint32_n("out_data",(uint32_t*)a, 10);
}

@@

        // initial value of sum.
        set_sum(1);

@@

        pthread_t acc_t, wpipe_t, rpipe_t;

#ifdef SW
        pthread_create(&acc_t,NULL,&accumulate_,NULL);
#endif
        pthread_create(&wpipe_t,NULL,&write_pipe_,(void*)data_in);
        pthread_create(&rpipe_t,NULL,&read_pipe_,(void*)data_out);

        pthread_join(wpipe_t,NULL);
        pthread_join(rpipe_t,NULL);

        fprintf(stdout,"from out_data, we read ");
        for(i=0; i < 10; i++)
          fprintf(stdout," %u ", data_out[i]);
        fprintf(stdout,"\n");
        fprintf(stdout,"final sum is %u\n", get_sum());

#ifdef SW
        pthread_cancel(acc_t);
#endif
}

\end{verbatim}
The test-bench sets an initial value for sum, and starts
three threads: one to write data to pipe in\_data, one
to read data from pipe out\_data, and one to run the
accumulate function.  After the last data is read back
from out\_data, the test-bench gets the value of sum
and prints it out.

Obviously, we would prefer to use the same test-bench
to verify that the VHDL system generated from ``prog.c'' 
functions correctly.  
The difference is that instead of using methods in iolib,
the test-bench now uses methods in SocketLib.  Further, the
VHDL is executed in a VHDL simulator; the simulator communicates
with the testbench using sockets.
The {\em ifdef's} in the test-bench and the system program
indicate the difference between the pure software version
of the system-test-bench combination and the hardware-software
version.

Look at the subdirectory ``example'' which contains the files ``prog.c'' and
``testbench.c''.  A Makefile and README are also present.  Its all quite
self-explanatory.

\end{document}

>>>>

Consider the following program (lets say it is in file ``prog.c''):
\begin{verbatim}
#include <stdlib.h>
#include <stdint.h>
#include <Pipes.h>
#include <stdio.h>

// note: initialized value..
uint32_t sum1 = 23;
uint32_t sum2 = 39;

// note: no problems with pointers :-)
uint32_t* tgt[2] = {&sum1, &sum2};

uint32_t get_sum(uint32_t idx)
{
  return(*(tgt[idx]));
}

void accumulate()
{
  int i = 0;
  while(1)
  {
     int nxt = read_uint32("in_data");
#ifdef SW
     printf("read %u\n", nxt);
#endif
     // ugly, but this is just a demo,
     // we are showing off.
     *(tgt[i])= ((*tgt[i]) + nxt);

     write_uint32("out_data",*(tgt[i]));
#ifdef SW
     printf("wrote %u\n", *(tgt[i]));
#endif
     i = 1 - i;
  }
}

@@

Now to test this program, we can write a test-bench
such as this one (lets call this file ``testbench.c'').
\begin{verbatim}
#include <pthread.h>
#include <signal.h>
#include <stdio.h>
#include <stdlib.h>
#include <stdint.h>

#include <Pipes.h>
#ifdef SW
#include <pipeHandler.h>
#include "prog.h"
#else
#include "vhdlCStubs.h"
#endif

@@

// The model of the hardware accumulate thread
// is necessary when building the SW testbench.
#ifdef SW
DEFINE_THREAD(accumulate)
#endif

@@

#ifdef SW
  init_pipe_handler();

  // register FIFO
  register_pipe("in_data",10,32,0);
  // register FIFO..
  register_pipe("out_data",10,32,0);

#endif

#ifndef SW
  // to set the initial value of sum.
  // in the hardware version, storage
  // variables are initialized by calling
  // this function (auto-generated by
  // the Aa linker AaLinkExtMem)
  global_storage_initializer_();
#endif

#ifdef SW
  // start the accumulate thread.
  // (this is necessary only in the SW model)
  PTHREAD_DECL(accumulate)
  PTHREAD_CREATE(accumulate)
#endif

@@

  // write 10 things to in_data, it has enough room..
  write_uint32_n("in_data",(uint32_t*)data_in, 10);

  // read back 10 things from out_data..
  read_uint32_n("out_data",(uint32_t*)data_out, 10);

  fprintf(stdout,"from out_data, we read ");
  for(i=0; i < 10; i++)
    fprintf(stdout," %u ", data_out[i]);
  fprintf(stdout,"\n");

  fprintf(stdout,"Sum 0 is %d\n",get_sum(0));
  fprintf(stdout,"Sum 1 is %d\n",get_sum(1));

#ifdef SW
  close_pipe_handler();
#endif
}
\end{verbatim}
In the SW case, the test-bench starts the accumulate thread,
writes data to the hardware, and reads back stuff from the
hardware.  In the non-SW case, there is no need to 
start the accumulate thread, since it exists in the 
hardware model.

Obviously, we would prefer to use the same test-bench
to verify that the VHDL system generated from ``prog.c'' 
functions correctly.  
The difference is that instead of using the pipeHandler,
the test-bench now uses methods in SocketLib.  Further, the
VHDL is executed in a VHDL simulator; the simulator communicates
with the testbench using sockets.
The {\em ifdef's} in the test-bench and the system program
indicate the difference between the pure software version
of the system-test-bench combination and the hardware-software
version.

The following Makefile builds a software-only testbench
executable, and also converts the system described in prog.c
to VHDL.  The same testbench can be used to test the VHDL
also.
\begin{verbatim}
# build software version of testbench (to check the "desired behaviour")
SOCKETLIB_INCLUDE=$(AHIR_RELEASE)/CtestBench/include
SOCKETLIB_LIB=$(AHIR_RELEASE)/CtestBench/lib
PIPEHANDLER_INCLUDE=$(AHIR_RELEASE)/pipeHandler/include
PIPEHANDLER_LIB=$(AHIR_RELEASE)/pipeHandler/lib
PTHREADUTILS_INCLUDE=$(AHIR_RELEASE)/pthreadUtils/include
VHDL_LIB=$(AHIR_RELEASE)/vhdl
VHDL_VHPI_LIB=$(AHIR_RELEASE)/CtestBench/vhdl
FUNCTIONLIB=$(AHIR_RELEASE)/functionLibrary/
SRC=./src
all: SW HW 
TOAA:c2llvmbc llvmbc2aa  aalink
TOVC:c2llvmbc llvmbc2aa  aalink aa2vc 
VC2VHDL: vc2vhdl  vhdlsim
AA2VHDLSIM: aa2vc  vc2vhdl  vhdlsim
TOVHDL:TOVC vc2vhdl

# llvm2aa opts: pipelined case, extract-do-while.
#LLVM2AAOPTS=--storageinit=true
LLVM2AAOPTS=-extract_do_while=true --storageinit=true -pipedepths=pipedepths.txt

PROGDEFS=

# the top-level modules
# -T specifies top-level daemon module
# -t specifies top-level slave module.
#
TOPMODULES=-T accumulate -t get_sum -t global_storage_initializer_

# compile with SW defined.
# note the use of IOLIB in building the testbench.
SW: $(SRC)/prog.c $(SRC)/prog.h $(SRC)/testbench.c 
	gcc -g -c -DSW $(PROGDEFS) -I$(PIPEHANDLER_INCLUDE)\
              -I$(FUNCTIONLIB)/include -I$(SRC) $(SRC)/prog.c
	gcc -g -c -DSW $(PROGDEFS) -I$(PIPEHANDLER_INCLUDE)\
              -I$(PTHREADUTILS_INCLUDE) -I$(SRC) $(SRC)/testbench.c
	gcc -g -o testbench_sw prog.o testbench.o\
              -L$(PIPEHANDLER_LIB) -lPipeHandler -lpthread

# five steps from C to vhdl simulator.
HW: c2llvmbc llvmbc2aa  aalink aa2vc  vc2vhdl  vhdlsim

AA2VHDL: aa2vc vc2vhdl vhdlsim

# C to llvm byte-code.. use clang.
c2llvmbc: $(SRC)/prog.c $(SRC)/prog.h
	clang -O3 -std=gnu89 $(PROGDEFS)  -I$(SOCKETLIB_INCLUDE)\
              -I$(FUNCTIONLIB)/include -emit-llvm -c $(SRC)/prog.c
	opt --indvars --loopsimplify prog.o -o prog.opt.o
	llvm-dis prog.opt.o

# llvm byte-code to Aa..
llvmbc2aa:  prog.opt.o 
	llvm2aa $(LLVM2AAOPTS)  prog.opt.o | vcFormat >  prog.aa

# Aa to vC
aalink: prog.aa 
	AaLinkExtMem prog.aa | vcFormat > prog.linked.aa
	AaOpt -B prog.linked.aa | vcFormat > prog.linked.opt.aa

aa2vc: prog.linked.opt.aa
	Aa2VC -O -C prog.linked.opt.aa | vcFormat > prog.vc

# vC to VHDL
vc2vhdl: prog.vc
	vc2vhdl -O -S 4 -I 2 -v -a -C -e ahir_system\
                 -w -s ghdl $(TOPMODULES) -f prog.vc 
	vhdlFormat < ahir_system_global_package.unformatted_vhdl\
                 > ahir_system_global_package.vhdl
	vhdlFormat < ahir_system.unformatted_vhdl\
                 > ahir_system.vhdl
	vhdlFormat < ahir_system_test_bench.unformatted_vhdl\
                 > ahir_system_test_bench.vhdl

# build testbench and ghdl executable
# note the use of SOCKETLIB in building the testbench.
vhdlsim: vhdlTb ghdlModel

vhdlTb: $(SRC)/testbench.c vhdlCStubs.h vhdlCStubs.c
	gcc -c vhdlCStubs.c  -I$(SRC)\
                -I./ -I$(SOCKETLIB_INCLUDE)
	gcc -c $(SRC)/testbench.c\
                -I$(PTHREADUTILS_INCLUDE) -I$(SRC)\
                -I./ -I$(SOCKETLIB_INCLUDE)
	gcc -o testbench_hw testbench.o vhdlCStubs.o\
                -L$(SOCKETLIB_LIB) -lSocketLib -lpthread

ghdlModel: ahir_system.vhdl ahir_system_test_bench.vhdl ahir_system_global_package.vhdl
	ghdl --clean
	ghdl --remove
	ghdl -i --work=GhdlLink  $(VHDL_LIB)/GhdlLink.vhdl
	ghdl -i --work=ahir  $(VHDL_LIB)/ahir.vhdl
	ghdl -i --work=aHiR_ieee_proposed\
                $(VHDL_LIB)/aHiR_ieee_proposed.vhdl
	ghdl -i --work=work ahir_system_global_package.vhdl
	ghdl -i --work=work ahir_system.vhdl
	ghdl -i --work=work ahir_system_test_bench.vhdl
	ghdl -m --work=work -Wl,-L$(SOCKETLIB_LIB) -Wl,-lVhpi ahir_system_test_bench

PHONY: all clean	
\end{verbatim}
To test the software, run testbench\_sw.  To verify the hardware (using
the VHDL simulator GHDL), start testbench\_sw in one shell, and then start ahir\_system\_test\_bench
in a different shell.
\end{document}
